How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)

### 2015-08-25 Manuscript.tex
Updates from ShareLaTeX

\title{How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Washington D.C.}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}
}
\maketitle
\begin{abstract}
 Business users now demand more insightful
 analytics. Instead of just predicting 
 some result, they also need tools that generate ``plans'';
 i.e. specific suggestions on  what to change  in order to
 improve on the predicted values. Hence, we propose the  XTREE planner and  show that
 it runs better than   three alternative methods. For this paper's    case studies,
 XTREE reduced
 the expected values of defects \& runtimes of some software projects to    
 45\% \& 61.5\%  (median) and  22\% \& 6\% (best case) of the initial values, respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} approach called XTREE for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREE
to reduce  the expected value of the
defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10};
and the runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.

The approach discussed here is novel since most data miners used in software analytics
comment on ``what is'' and not ``what to do''. For example, a clustering
algorithm might find clumps of similar examples, but does not comment on the minimal
change(s) required to move from clump to clump.

The contributions of this paper are (1)  
   XTREE   (which is a new algorithm) and (2) an evaluation strategy for   planners.
XTREE     works  on tables of data with weighted class values indicating what rows are ``good'' and what rows are ``bad''.
In the case studies of this paper  our evaluations shows
 XTREE performing significantly ``better'' than   the planners we proposed in  prior   papers~\cite{me12c,krishna15},
 where ``better'' means:
\bi
\item {\em Effective:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  in expected values of the results should  be  observed. 
\item {\em Succinct:} i.e. given two plans with the same effectiveness, we prefer the shorter one.
Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
\item {\em Surprising}; i.e. we want the planners to (occasionally) tell us things we do not know already (otherwise,
there would be no point to using the planner).
\ei 
The rest of this paper is describes our data, our planners, and the experiment that
ranks XTREE against alternate approaches.  This is followed by notes on related work
and validity.
Note that, to allow for reproducibility, all scripts and data used in this 
study are available on-line at github.com/ai-se/XTREE.

@@

@@

\begin{figure}[!t]
  \includegraphics[width=\linewidth]{figs/twodee.eps} 
 % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{Training examples (in gray), test examples (in red) and 
changed examples generated by a planner (in green).
Displayed on the first two components of a PCA analysis of all points. 
The green (changed) examples are closer to the  training
date than the original test data
i.e. predictors trained on the training examples and
applied to the test examples should
work just as well, or better, on the changed test examples.}\label{fig:howxy}
\end{figure}

@@

{\em Assumption4:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
the algorithm used for the initial clustering. Unsupervised methods like WHERE
execute without knowledge of the target class.  {\em Supervised} methods, on the other hand, assume that it is useful to also reflect on the target class.

{\em Details4:} 
XTREE uses a supervised   decision tree algorithm of \fig{where}.D to divide the data.
Next, XTREE builds plans from the branches of the decision trees. It does so by asking three questions:
\be
\item
What {\em current} branch does a test case fall in?
\item What {\em desired} branch would the test case want to move to?
\item What are the {\em deltas} between current and desired? 
\ee
The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{xtrees_bare}.

@@

An  interesting feature of \fig{changed} was that fewer things 
were changed in the config data sets {\em BDBJ} than in the defect data set
{\em lucene}. 
In turns out that this holds true across nearly all our data sets.
\fig{types} summarizes all the change frequencies for all data sets. As with  \fig{changed},
there are fewer changed features in the config data than in the defect prediction data. 
One explanation for that is the nature of the features: the
defect data sets has continuous
features while the 
config data has binary independent features  (where so setting was turned ``on'' or ``off''). When exploring these different data types, it is possible to find more
\bi
\item
``Fentle slopes'' the lead to small changes
in  continuous space;
\item
 ``Sharp cliffs'' that lead to major change in the discrete
space. 
\ei
Hence, our planners can make fewer larger changes in the config data.
 For example, Method3 (BIC) and Method1 (CD) make far fewer changes in   data sets with discrete features (the Siegmund data) than in data with continuous features (the Jureczko data).

@@

Such a plan was seen in the {\em lucene} results of \fig{changed}. Recall the standard advice for
OO systems: build classes that are internally cohesive with low coupling to other parts of the system.  
\subsubsection{Summary}
In summary, we  have two reasons to endorse XTREE. Firstly, it is usually far more effective than other
methods. Secondly, even when we jump from continuous to discrete features, XTREE's plans consistently
use just a small portion of the features.

In terms of the assumptions that motivated our different methods, we note that 
{\em Assumption4} seems most powerful; i.e. for the purposes of planning,
it is better to reflect on the target class.

@@

\subsection{  Sampling bias} threatens any data mining experiment; i.e., what matters
there may not be true here. For example, the data sets used here comes from two sources
(Seigmund et al. and Jureczko et al.) and any biases in their selection procedures
threaten the validity of these results. 
That said,
the best we can do is define our methods and publicize our data and code so that other researchers can
try to repeat our results and, perhaps, point out a previously unknown bias
in our analysis. Hopefully, other researchers will emulate our methods in
order to repeat, refute, or improve our results.

\subsection{  Evaluation bias} Another threat to validity of this work is our use
of predictors learned on the training data to assess the impact of our planners.
This issues was discussed in detail in \tion{trust}.

@@

For example, the COCOMO model from University of 
Southern California estimates software development time using a combination of industrial data and some domain expertise
from its author (Barry Boehm). \fig{coc} shows results from using COCOMO 
as an oracle to assess our planning methods. In this experiment, random projects were used as COCOMO
inputs. This generated an development time estimate for each example which we passed to our four methods.
40 times, we let those methods propose changes to those projects. 
For assessment purposes, the changes projects were then feed back to the COCOMO
oracle.

As shown in \fig{coc}, those results endorsed the conclusions of the rest
of this paper; i.e. that XTREE's supervised methods are best for generating plans on how to change
example projects.

\newpage
\section{Related work}

@@

\section{Conclusion}

works for more than defects and tunrimes

This main contribution of this paper was to  define an evaluation strategy for this kind of reasoning, and to show that (using that strategy) we
can (a)~select useful planning methods while (b)~culling clearly
inferior planning methods. 
We apply this strategy to XTREE and three alternate planning methods to find:
\bi
\item Two methods that are clearly inferior;
\item One method that proposes complex changes, which work as well as XTREE;
\item And XTREE, that proposes very simple changes.
\ei
Of course, there are many more methods for generating plans and
no   one paper can survey them all.
But the goal of this paper is not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
Our hope is that other   researchers apply this strategy  to repeat and/or improve
and/or refute our results. To that end, all our materials and data sets are available on-line, see
github.com/ai-se/XTREE.

>>>>

\title{How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Washington D.C.}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}
}
\maketitle
\begin{abstract}
 Business users now demand more insightful
 analytics. Instead of just predicting 
 some result, they also need tools that generate ``plans'';
 i.e. specific suggestions on  what to change  in order to
 improve on the predicted values. Hence, we propose the  XTREE planner and  show that
 it runs better than   three alternative methods. For this paper's    case studies,
 XTREE reduced
 the expected values of defects \& runtimes of some software projects to    
 45\% \& 61.5\%  (median) and  22\% \& 6\% (best case) of the initial values, respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} method called XTREE for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREE
to reduce  the expected value of the
defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10};
and the runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.

The contributions of this paper are (1)  
    the new XTREE    algorithm and (2) an evaluation strategy for   planners.
In the case studies of this paper  our evaluations shows
 XTREE performing significantly ``better'' than    planners  proposed in our prior   work~\cite{me12c,krishna15},
 where ``better'' means:
\bi
\item {\em Effective:}    i.e.  if  a plan is   applied  then  some  statistically significant  change  in expected values of the results will  be  observed. 
\item {\em Succinct:} i.e. given two plans with the same effectiveness, we prefer the shorter one.
Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
\item {\em Surprising}; i.e. we want the planners to (occasionally) tell us things we do not know already (otherwise,
there would be no point to using the planner).
\ei 
The rest of this paper is describes our data, our planners, and the experiment that
ranks XTREE against alternate approaches.  This is followed by notes on related work
and validity.
Note that, to allow for reproducibility, all scripts and data used in this 
study are available on-line at github.com/ai-se/XTREE.

@@

\subsection{What is a ``Plan''?}

Our planners   use tables of data with independent features and one dependent feature, called the 
``class''. Classes are weighted such that they indicate  what rows are ``good'' and what rows are ``bad''. Plans  change   a row such that it is more like they 
will be more similar to ``good'' than ``bad''.
Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
  adjust   feature $Z_j$:
\[
\forall \delta_j \in \Delta :  Z_j = Z_j + \delta_j
\]
For example, to simplify a  large bug-prone  method, our planners might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).

Note that we make no assumption that a plan mentions every features
(so plan1 can be  more succinct than plan2 when plan1  mentions fewer features than
plan2).

@@

\begin{figure}[!t]
  \includegraphics[width=\linewidth]{figs/twodee.eps} 
 % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{Gray, red, green show (1) training examples, (2) test examples and 
  (3) tests that have been altered by planners.
  This figure uses axes generated from the first two components of a PCA analysis of all points. 
}\label{fig:howxy}
\end{figure}

@@

{\em Assumption4:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
the clustering algorithm (WHERE) that  
executes without knowledge of the target class.  {\em Supervised} methods, on the other hand, assume that it is useful to also reflect on the target class.

{\em Details4:} 
XTREE uses a supervised   decision tree algorithm of \fig{where}.D to divide the data.
Next, XTREE builds plans from the branches of the decision trees using the code of \fig{xtrees_bare}.
That code asks three questions, the last of which returns the plan:
\be
\item
What {\em current} branch does a test case fall in?
\item What {\em desired} branch would the test case want to move to?
\item What are the {\em deltas} between current and desired? 
\ee

@@

An  interesting feature of \fig{changed} was that fewer things 
were changed in the config data sets {\em BDBJ} than in the defect data set
{\em lucene}. 
In turns out that this holds true across nearly all our data sets.
\fig{types} summarizes all the change frequencies for all data sets. As with  \fig{changed},
there are fewer changed features in the config data than in the defect prediction data. 
One explanation for that is the nature of the features: the
defect data sets has continuous
features while the 
config data has binary independent features  (where so setting was turned ``on'' or ``off''). When exploring these different data types, it is possible to find more
\bi
\item
``Gentle slopes'' the lead to small changes
in  continuous space;
\item
 ``Sharp cliffs'' that lead to major change in the discrete
space. 
\ei
Hence, our planners can make fewer larger changes in the config data.
 For example, Method3 (BIC) and Method1 (CD) make far fewer changes in   data sets with discrete features (the Siegmund data) than in data with continuous features (the Jureczko data).

@@

Such a surprising plan can be  seen in the {\em lucene} results. Recall the standard advice for
OO systems: build classes that are internally cohesive with low coupling to other parts of the system~\cite{Dhama199565}. We can assess the relevance of this advice
to specific projects by checking how often a planner changes the
 coupling-related features:
\bi
\item {\em ca}:   afferent couplings  =  how many other classes use the specific
				class;
\item {\em  	ce}:  efferent couplings =  how many other classes is used by the
				specific class. 
\item {\em cbm}: coupling between methods =  total number of new/redefined methods
				to which all the inherited methods are coupled
\item	{\em cbo}:  coupling between objects = a value that increases when the methods of one
				class access services of another.

\item {\em ic}:   inheritance coupling =  number of parent classes to which a given
				class is coupled (includes counts of methods and variables inherited)
\ei
For the  {\em lucene} XTREE results  of  \fig{changed},
the most frequent change was to alter the lines of code in a class (see the tallest
red historgram in that figure on the {\em loc}, or lines of code). 
Looking at the logs of our planner, we can see
that XTREE's proposed changed is to {\em reduce} the size of a class.
The only way to do that,  while keeping the  same functionality, it is create a network
of smaller classes that interact to produce that functionality.
That is, we would need to {\em increase} the coupling of those classes to achieve XTREE's plan.

In theory, increasing coupling between classes complicates and confuses a class design.
But the {\em lucene} XTREE results  of  \fig{changed} rarely proposes changing   the
coupling features {\em ca, ce, cbm, ic} (in fact, XTREE never proposes any change to {\em ic}).

The only coupling issue that XTREE   usually adds to its plans is {\em cbo} (which appears 55\% of the time
in \fig{changed}). But note  that this is {\em object} coupling measure, not class coupling. So here
XTREE is warning against, say, some factory class generating a large
community of agents, all of the same class, who
co-ordinate on some task. This is a different issue to the class redesign issue that would
be triggered by altering {\em loc}}.

In summary, XTREE satisfies that criteria that, sometimes, it produces surprising plans.
At least for the {\em lucense} data set, we can see advice that recommends {\em increasing coupling}
to reduce defects.

@@

\subsection{  Sampling bias} 
Sampling bias threatens any data mining experiment; i.e., what matters
there may not be true here. For example, the data sets used here comes from two sources
(Seigmund et al. and Jureczko et al.) and any biases in their selection procedures
threaten the validity of these results. 
That said,
the best we can do is define our methods and publicize our data and code so that other researchers can
try to repeat our results and, perhaps, point out a previously unknown bias
in our analysis. Hopefully, other researchers will emulate our methods in
order to repeat, refute, or improve our results.

\subsection{  Evaluation Bias} Another threat to validity of this work is our use
of predictors learned on the training data to assess the impact of our planners.
This issues was discussed in detail in \tion{trust}.

@@

For example, the COCOMO model from University of 
Southern California estimates software development time using a combination of industrial data and some domain expertise
from its author (Barry Boehm). \fig{coc} shows results from using COCOMO 
as an oracle to assess our planning methods. In this experiment, random projects were used as COCOMO
inputs. This generated an development time estimate for each example, which we passed to our four methods.
40 times, we let those methods propose changes to those projects. 
For assessment purposes, the changes projects were then feed back to the COCOMO
oracle.

Using this approach, it is possible to assess the value of a plan by measuring its
effectiveness with respect to some ground truth (in this case, the COCOMO model).
As shown in \fig{coc}, XTREE passes this assessment.
Those results from the COCOMO model endorsed the conclusions of the rest
of this paper; i.e. compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change
example projects.

\section{Related work}

@@

\section{Summary}

The planner proposed in this paper propose changes to software project details in order to improve the expected
value of the performance scores of that part of the project.
To evaluate these planners,
data miners can be used to build oracles to assess planners.
Such planners should be {\em trust-increasing}; i.e. they propose changes that generate
changed examples that are closer to the training data of the data miner.
One caveat here is that the evaluations we can make on the planner are only as good as the predictive
performance of the data miner. Hence, if domain data does not support satisfactory predictors, then
planning in that domain cannot be evaluated.

Four planners were assessed here for the tasks of reducing defects and runtimes. 
Three of the methods come from our prior publications~\cite{me12c,krishna15}, and the conclusion of this
paper is that a novel fourth method clearly out-performs the other three
(measured in terms of {\em effectiveness, succinctness, and surprise}).
The reason for the superior performance of our new method is clear: when planning how to change
to different class methods, it is best to use a learner that reflects over those classes
as it divides the data.

Of course, there are many more methods for generating plans and
no   one paper can survey them all.
But the goal of this paper is not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
Our hope is that other   researchers apply this strategy  to repeat and/or improve
and/or refute our results.
